### 2022-08-08 main/paper3.tex
Add files via upload

Can the stock market be predicted by analyzing social media? Recent developments in machine learning and the growing quantities of available text data from online news, social media and annual reports have triggered intensive research in finance. In their pioneering paper, \citet{antweiler2004all} compute a bullishness measure out of 1.5 million messages posted on Yahoo! Finance and Raging Bull and finds that stock messages help predict market volatility. Their results clearly reject the hypothesis that all that talk is just noise. They show that there is financially relevant information present in social media. In a similar vein, \citet{tet_07} constructs a measure of media pessimism from a Wall Street Journal column and finds that it predicts downward pressure on market prices.

Most of the previous financial studies of social media rely on pre-defined or manually annotated sentiment dictionaries. Such approaches are limited in various ways. How to create a sentiment classifier that understands the vocabulary of the messages posted by the investors? For instance, ``bull" is an animal in everyday language but it is someone optimistic in the financial jargon. \citet{loughran2011liability} create a word list, which helps classify tone in a financial document. However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing. For instance, the word ``goooooood" would not be recognized by the model if it is not corrected into ``good" first. On the other hand, manually annotating and validating dictionaries is not a scalable approach to handling social media content.

Our paper overcomes these limitations. We develop a machine learning algorithm to classify the sentiment of a large sample of StockTwits messages as bullish, bearish, or neutral. The sample consists of all messages referring to US and Canadian stocks, including ETFs and other types of securities available on CRSP/Compustat, from January 2010 to March 2020. We train our machine learning classifier on the set of all user sentiment-labeled messages, which constitute about one third of the sample. We then classify the sentiment of all remaining messages. Our method scales and performs very well. It achieves an out-of-sample accuracy of 85.9\%, which compares well to the anecdotal 80 to 85\% probability that human annotators agree on the sentiment of a document, see, e.g., \citet{wilson2005recognizing} and \citet{chen2020large}. As a side product, we generate a vocabulary of one million investor sentiment-labeled terms consisting of up to three words that frequently appear in StockTwits messages.

@@

As a technical byproduct, we develop a sentiment classifier of micro-blogs for imbalanced data. This addresses the stylized fact that bloggers post more bullish than bearish-labeled messages. In our sample, the ratio is five to one. On the other hand, we find that not all messages carry a substantial stock market relevant sentiment. Rather than re-sampling from the underrepresented bearish class, we thus introduce an auxiliary neutral class. We then run two independent binary classifiers. The first (second) classifies messages as bullish versus non-bullish (bearish versus non-bearish). We aggregate the two binary outcomes and classify a message as bullish (bearish) for the concordant combination bullish/non-bearish (non-bullish/bearish), and neutral otherwise. This approach is very simple and efficient, and eliminates the class imbalance bias at the same time. It builds on any traditional binary classifiers. We use logistic regression on Term Frequency-Inverse Document Frequency (TFIDF)-vectorized messages. TFIDF is a weighting scheme gauging the importance of a word in a document.

Our paper contributes to the growing literature on machine learning classification of social media and its interaction with the stock market. Most previous financial studies use Twitter as their primary source of data. Twitter has the advantage of being used by a wide range of people across the world and a few influencers can attract the attention of many investors. In 2013, following a meeting with Tim Cook (Apple CEO), Carl Icahn tweeted that he bought a large position in Apple and believed that the company is extremely undervalued. This bullish tweet caused the market capitalization of Apple to jump by \$12 billion. In 2019, JPMorgan has created the Volfefe Index to track Donald Trump's tweets impact on the stock market. However, it is more difficult to disentangle noise from relevant tweets in Twitter than in other more focused social media. Results from \citet{Ghoshal} show that StockTwits is significantly more informative than Twitter data. This is not surprising as StockTwits is a finance-focused platform whereas Twitter also captures irrelevant opinions on a wide range of non-finance related matters.

Our paper is the first that analyzes the predictive power of StockTwits messages on stock returns unconditionally and around specific events. \citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day. However, his classifier is based on a dictionary consisting of 8 thousand manually validated and modified terms, which limits its scalability. \citet{renault2020sentiment} uses larger data sets and compares various classifiers, including machine learning.

@@

\citet{ke_kel_xiu_20} extract sentiment from news articles on the Dow Jones Newswires. They train a  sentiment score directly on returns. In contrast, we use user sentiment-labeled StockTwits messages as training and validation set for our sentiment classifier.

Our paper also contributes to the EMH literature by gauging how cumulative average abnormal returns and abnormal polarities behave around sudden peaks of message activity.

The remainder of the paper is structured as follows. Section \ref{S:data} discusses StockTwits and stock market data. Section \ref{S:classification} develops our sentiment classifier based on TFIDF vectorization. Section \ref{S:polarity} introduces the sentiment polarity measure and relates it to stock returns. Section \ref{S:event studies} contains the event study. Section \ref{S:portfolios} discusses the sentiment-sorted portfolio performance. Section \ref{S:conclusion} concludes. The appendix contains additional statistics and background material.

@@

The reasons for using StockTwits and not other social media data for financial studies are at least threefold. First, a major challenge in applying natural language processing is the creation of an appropriate labeled vocabulary. \citet{loughran2011barron} show that it is essential to have a specific vocabulary to interpret finance documents (i.e., many words have a different meaning in finance than in traditional English, such as “bear trap”). In addition to that, social media slang is an additional layer of language complexity. To this extent, the functionality to self-tag bullish and bearish messages that StockTwits is extremely valuable as it allows the creation of a specific labeled vocabulary out of labeled messages. We are not aware of any other social media platform in finance offering this functionality. Second, text data from StockTwits is more reliable and less noisy than from general purpose platforms, such as Twitter, because messages focus on finance and economics matters only. Micro-bloggers have incentives to post valuable information in order to maintain or increase mentions and retweets, and thus have a greater share of voice in the forum (\citet{sprenger2014tweets}). On the other hand, StockTwits messages might be biased and subject to malicious users that try to manipulate the market. However, market manipulations likely happen only rarely as the SEC closely monitors potential influencers to prevent any market abuse. Third, extracting data from StockTwits is easy because of its API. StockTwits' API is designed to query the database to download messages via JSON requests. We provide a short tutorial in Appendix~\ref{app_tut}.

@@

Every StockTwits message includes eight features: (1) the reference ticker(s), (2) a timestamp, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who posted the message, (7) the number of messages published by the user who posted the message, and (8) the number of followers of the user who posted the message. Our sentiment analyses builds on the first five features. The last three provide additional information on the network structure, which we briefly discuss in Appendix \ref{app_userstats}.

@@

Left plot of Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on StockTwits. SPY, a large ETF that tracks the S\&P 500 stock market index, is the most discussed ticker, followed by Apple and other big tickers. The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms. Right graph shows a histogram of the number of messages per ticker. The x-axis is log-scaled because due to extreme values the distribution is highly skewed.

@@

Left plot of Figure \ref{fig:classif_time} shows the proportions of user sentiment-labeled messages across time. In the early years of the platform, most messages were unlabeled, presumably because users were not familiar with the sentiment label yet. Albeit the proportion of unlabeled messages monotonically declines over the years, almost 60\% of the more recent messages are still unlabeled. Overall, around 30 million messages are user-labeled and 60 million messages are unlabeled. We conjecture that by far not all unlabeled messages reflect market neutral opinions. Indeed, the right plot of Figure \ref{fig:classif_time} reveals that a substantial part of user-unlabeled messages is machine learning classified as bullish or bearish. Hence these user-unlabeled messages contain indeed market relevant information, which we are able to capture by our algorithm.

@@

    \mycaption{Message classification}{Left plot shows the proportions of user-labeled messages: bullish (green), bearish (red), and unlabeled (gray) across time. Right plot shows the proportions of machine learning classified messages: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray) across time. Proportions are aggregated monthly.}
    \label{fig:classif_time}

@@

Among the user-labeled messages we find five times more bullish than bearish ones. This ratio indicates that investors are on average optimistic about the market, which is consistent with findings in the literature, e.g., \citet{renault2017intraday}. Such an imbalance is a well-known issue in machine learning classification as it creates a bias towards the over-represented class, see \citet{chawla2004special}. There are various ways to tackle class imbalance. An all-purpose standard approach in machine learning is to over-sample the minority class, which consists of randomly re-sampling from the minority class and thus artificially re-balance the class sizes in the data. We use a different approach, which is tailored for our setup. As not every message carries a substantial stock market relevant sentiment, we deviate from the bullish--bearish dichotomy. Instead, we create an auxiliary neutral sentiment class to account of messages that do not take a clear stand. See Appendix \ref{app_exampleclass} for examples of such neutral messages.

@@

Right plot of Figure \ref{fig:classif_time} shows the proportions of our machine learning classifications across time. Percentages of bearish (user-labeled and classified as bearish) and bullish (user-labeled and classified as bearish) messages are stable over time, suggesting that our classification method is robust. Even if most messages were not user-labeled in the early years of the platform, as seen in the left plot of Figure \ref{fig:classif_time}, we are now able to classify the sentiment of all messages in the sample, including a neutral class. Consistent with the over-representation of bullish messages observed in the user-labeled messages, there are substantially more messages classified as bullish than bearish. Examples of classified messages are given in Appendix \ref{app_exampleclass}.

@@

We next aggregate message sentiments on a daily ticker-level and across the market. Thereto, we denote by $C_{i,t,j}=1,0,-1$ for bullish, neutral, bearish, respectively, the sentiment of the $j$th message about ticker $i$ on day $t$.\footnote{We follow the close-to-close convention. First, we remove all non-business days from the sample, whereby messages posted on non-business days count towards the next business day. ``Day $t$'' then stands for the time interval from 4pm on the previous business day $t-1$ to 4pm on business day $t$. This convention is consistent with the stock return data, which are close-to-close, and thus avoids any look-ahead bias of our sentiment polarity.} We follow \citet{ranco2015effects} and define the sentiment polarity of ticker $i$ on day $t$ as
\begin{equation}
    P_{i,t} = \dfrac{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} - \textbf{1}_{C_{i,t,j} = -1})}{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} + \textbf{1}_{C_{i,t,j} = -1})},
\end{equation}
where $V_{i,t}$ denotes the number of messages about ticker $i$ on day $t$.\footnote{If $V_{i,t}=0$ then we set $P_{i,t} =0$.}

@@

For the following time-series analysis and event study we restrict our sample. There are two reasons for doing so. First, we keep computational cost at a reasonable level. Second, and more importantly, the time series of ticker-individual polarities exhibit spikes and are too noisy if the daily message volumes $V_{i,t}$ are too small. We therefore exclude from now on tickers whose median of daily message volume is less than 50. Our reduced sample contains 19 tickers, covering a range of security types, sectors, and market capitalization. We refer to Appendix \ref{app_coverage} for more details on the coverage, including summary statistics and the list of tickers covered.

@@

Event studies constitute a statistical method widely used in financial econometrics, see, e.g., \citet{mackinlay1997event}. In general, they are used to measure the effect of events on the market value of stocks. Well-known applications of event studies include the testing of various forms of the efficient market hypothesis (EMH) (see \citet{fama1969} and \citet{fama1991efficient}).

@@

Next we define the type of the event as either bullish, neutral or bearish. We use the abnormal polarity $AP_{i,t}$ of the event date to assess how on average investors perceive the event. Figure \ref{fig:abnpol_all} shows the distribution of abnormal polarities on event dates. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{abnpol_distr_all.JPG}
    \mycaption{Empirical distribution of abnormal polarities on event dates}{Red dashed lines show the one-third and two-third percentiles.}
    \label{fig:abnpol_all}
\end{figure}
We chose to use the one-third (-0.03) and two-third percentile (0.07) of the distribution of abnormal polarities as thresholds for the type of the event. We define the type of the event for ticker $i$ at $t$ as
\begin{equation}
Type_{i,t}= \begin{cases}
      Bullish  & \text{if}\ \ AP_{i,t}> 0.07, \\
      Neutral  & \text{if}\ \ AP_{i,t} \in [-0.03, 0.07], \\
      Bearish  & \text{if}\ \ AP_{i,t} < -0.03. \\
\end{cases}    
\end{equation}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{events_across_time.JPG}
    \mycaption{Number of events of each type across time}{Green area shows bullish events, gray area shows neutral events and red area shows bearish events. Numbers are aggregated monthly.}
    \label{fig:events}
\end{figure}

As an illustration, Figure \ref{fig:activityaapl} shows for Apple the time-series of message volume and the the corresponding events. Between January 2011 and March 2020, our algorithm identified 73 events for Apple. 
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.75\textwidth]{activity_aapl.JPG}
    \mycaption{Daily message volume for Apple}{Events are days with an unusual high number of messages. Green upper-triangles show bullish events, gray circles are neutral events and red down-triangles represent bearish events.}
    \label{fig:activityaapl}
\end{figure}
What are these events? Remarkably, we capture a variety of corporate events and disclosures. Earning announcements constitute about half of the events. Other events include Apple Keynotes (presentations that Apple gives to the press, often presenting new products), or CEO letters addressed to investors. Table \ref{tab:event_aapl} lists a few examples.

@@

Left plot of Figure \ref{fig:caarcaap} shows the CAAR around the events. This plot is consistent with \citet{mackinlay1997event}. It shows that CAAR related to bearish (bullish) events exhibits a significant downward (upward) jump at the event date, respectively. These jumps are followed by a flat CAAR during the 20 days after the event. Interestingly, there is a systematic shift in the CAAR already 1 day before the event. However, this shift is relatively small compared to the jump on the event day: one day before the event, the bullish (bearish) CAAR equals 0.019 (-0.023). The CAAR related to the neutral events exhibits a slight upward shift around the event date but it fades away after a few days. The CAAR related to bearish events shifts already a few days before the event but this shift is not statistically significant. This is in line with Figure \ref{fig:boxplotcaar}, which shows that the CAR distributions prior to the events are not significantly different from zero. This is confirmed by the Mann-Whitney U-tests shown in Table \ref{utest_res}. CAR has no predictive power on the type of the event: five days before an event, the median of the CAR distribution of the bullish events is not statistically different from the median of the neutral events. Same holds for the bearish events.

@@

%\clearpage
\subsection{Abnormal Polarity}
How does sentiment polarity behave around events? Similar to the above, we use as benchmark model a one-year rolling window regression of the the daily polarity of ticker $i$ on the daily market polarity defined in \eqref{marketpoleq}, 
\begin{equation}
    P_{i,t}= \alpha_i^P  + \beta_i^P \cdot P_t^M  + \epsilon_{i,t},
\end{equation}
which gives the one-year rolling estimates $\hat{\alpha}_i^P$ and $\hat{\beta}_i^P$. This implies the abnormal polarity 
\begin{equation}
      AP_{i,t} = P_{i,t} - \left(\hat{\alpha_i}^P + \hat{\beta_i}^P \cdot P_t^M\right).
\end{equation}
We define the cumulative abnormal polarity (CAP) around a ticker $i$ event $\tau$ as 
\begin{equation}\label{eq_CAPtaut}
    CAP_i(\tau,t) = \sum_{s=- 20}^{t} AP_{i,\tau + s},
\end{equation}     
and the cumulative average abnormal polarities (CAAP) across all $N=1131$ events as 
\begin{equation}
       CAAP(t) = \dfrac{1}{N} \sum_{j=1}^N CAP_{i_j}(\tau_j,t).
\end{equation}

Right plot of Figure \ref{fig:caarcaap} shows the CAAP around the events. There are two main findings. First, in contrast to CAAR, the CAAP for bullish and bearish events is not constant after the event date, suggesting that users' sentiments about stocks tend to be biased towards recent past events. A possible explanation is that users might still post bullish (bearish) messages about a bullish (bearish) event during several days after the event. This is in contrast to the returns that immediately normalize after the event. Second, and more interestingly, the CAAP for bullish and bearish events shifts several days earlier than the CAAR. This indicates that investors are on average able to anticipate the type of an event in the near future. However, this sentiment only manifests through the social media, but not through abnormal returns.

@@

Based on these signals, we now construct reset CAP-sorted portfolios. Formally, we define the ticker sets $I^{bull}_t=\{ i\mid CAP_{i,t}^{(R)} > U_t(x)\}$ and $I^{bear}_t=\{ i\mid CAP_{i,t}^{(R)} < L_t(x)\}$. At the close of any day $t$, we form the equally weighted bullish (bearish) portfolio consisting of tickers in $I^{bull}_t$ (in $I^{bear}_t$), and realize the 1-day returns. If any of the index sets is empty, we set the corresponding return to zero. Top-left plot of Figure \ref{fig:pf} shows the cumulative log returns of bullish and bearish portfolios as well as the S\&P500. Overall, the portfolio performance is consistent with our approach: the bullish (bearish) portfolio outperforms (under-performs) the market. Remarkably, the upward (downward) steps suggests that our portfolio strategy succeeds to take the right positions just before an event. The remaining plots of Figure \ref{fig:pf} show the number of positions across time of our portfolios. Most of the returns are earned with portfolios consisting of very few tickers. This is a result of our market timing and stock picking strategy: we only invest in the top/bottom percentiles of CAP, whenever our signal is strong enough.

>>>>

Can the stock market be predicted by analyzing social media? Recent developments in machine learning and the growing quantities of available text data from online news, social media and annual reports have triggered intensive research in finance. In their pioneering paper, \citet{antweiler2004all} compute a bullishness measure out of 1.5 million messages posted on Yahoo! Finance and Raging Bull and find that stock messages help predict market volatility. Their results clearly reject the hypothesis that all that talk is just noise. They show that there is financially relevant information present in social media. In a similar vein, \citet{tet_07} constructs a measure of media pessimism from a Wall Street Journal column and finds that it predicts downward pressure on market prices.

Most of the previous financial studies of social media rely on pre-defined or manually annotated sentiment dictionaries. Such approaches are limited in various ways. How to create a sentiment classifier that understands the vocabulary of the messages posted by the investors? For instance, ``bull" is an animal in everyday language but refers to someone optimistic in the financial jargon. \citet{loughran2011liability} create a words list, which helps classify tone in a financial document. However, this might not be sufficient in the context of social media because messages posted present many typos, abbreviations and slang, so one needs to have an additional layer of data preprocessing. For instance, the word ``goooooood" would not be recognized by the model if it is not corrected into ``good" first. On the other hand, manually annotating and validating dictionaries is not a scalable approach to handling social media content.

This chapter overcomes these limitations. We develop a machine learning algorithm to classify the sentiment of a large sample of StockTwits messages as bullish, bearish, or neutral. The sample consists of all messages referring to US and Canadian stocks, including ETFs and other types of securities available on CRSP/Compustat, from January 2010 to March 2020. We train our machine learning classifier on the set of all user sentiment-labeled messages, which constitute about one third of the sample. We then classify the sentiment of all remaining messages. Our method scales and performs very well. It achieves an out-of-sample accuracy of 85.9\%, which compares well to the anecdotal 80 to 85\% probability that human annotators agree on the sentiment of a document (see, e.g., \citet{wilson2005recognizing} and \citet{chen2020large}). As a side product, we generate a vocabulary of one million investor sentiment-labeled terms consisting of up to three words that frequently appear in StockTwits messages.

@@

As a technical byproduct, we develop a sentiment classifier of micro-blogs for imbalanced data. This addresses the stylized fact that bloggers post more bullish than bearish-labeled messages. In our sample, the ratio is five to one. On the other hand, we find that not all messages carry a substantial stock market relevant sentiment. Rather than re-sampling from the underrepresented bearish class, we thus introduce an auxiliary neutral class. We then run two independent binary classifiers. The first (second) classifies messages as bullish versus non-bullish (bearish versus non-bearish). We aggregate the two binary outcomes and classify a message as bullish (bearish) for the concordant combination bullish/non-bearish (non-bullish/bearish), and neutral otherwise. This approach is very simple and efficient, and eliminates the class imbalance bias at the same time. It builds on any traditional binary classifier. We use logistic regression on Term Frequency-Inverse Document Frequency (TFIDF)-vectorized messages. TFIDF is a weighting scheme gauging the importance of a word in a document.

This chapter contributes to the growing literature on machine learning classification of social media and its interaction with the stock market. Most previous financial studies use Twitter as their primary source of data. Twitter has the advantage of being used by a wide range of people across the world and a few influencers can attract the attention of many investors. In 2013, following a meeting with Tim Cook (Apple CEO), Carl Icahn tweeted that he bought a large position in Apple and believed that the company is extremely undervalued. This bullish tweet caused the market capitalization of Apple to jump by \$12 billion. In 2019, JPMorgan has created the Volfefe Index to track Donald Trump's tweets impact on the stock market. However, it is more difficult to disentangle noise from relevant tweets in Twitter than in other more focused social media. Results from \citet{Ghoshal} show that StockTwits is significantly more informative than Twitter data. This is not surprising as StockTwits is a finance-focused platform whereas Twitter also captures irrelevant opinions on a wide range of non-finance related matters.

This chapter is the first work that analyzes the predictive power of StockTwits messages on stock returns unconditionally and around specific events. \citet{renault2017intraday} builds an intraday investor sentiment indicator using messages and finds that the change in investor sentiment of the first half-hour of a trading day helps forecast the last half-hour market return of that trading day. However, his classifier is based on a dictionary consisting of 8 thousand manually validated and modified terms, which limits its scalability. \citet{renault2020sentiment} uses larger data sets and compares various classifiers, including machine learning.

@@

\citet{ke_kel_xiu_20} extract sentiment from news articles on the Dow Jones Newswires. They train a  sentiment score directly on returns. In contrast, we use user sentiment-labeled StockTwits messages as the training and validation set for our sentiment classifier.

This chapter also contributes to the EMH literature by gauging how cumulative average abnormal returns and abnormal polarities behave around sudden peaks of message activity.

The remainder of the chapter is structured as follows. Section \ref{S:data} discusses StockTwits and stock market data. Section \ref{S:classification} develops our sentiment classifier based on TFIDF vectorization. Section \ref{S:polarity} introduces the sentiment polarity measure and relates it to stock returns. Section \ref{S:event studies} contains the event study. Section \ref{S:portfolios} discusses the sentiment-sorted portfolio performance. Section \ref{S:conclusion} concludes. The appendix contains additional statistics and background material.

@@

The reasons for using StockTwits and not other social media data for financial studies are at least threefold. First, a major challenge in applying natural language processing is the creation of an appropriate labeled vocabulary. \citet{loughran2011barron} show that it is essential to have a specific vocabulary to interpret finance documents (i.e., many words have a different meaning in finance than in traditional English, such as “bear trap”). In addition to that, social media slang is an additional layer of language complexity. To this extent, the functionality to self-tag bullish and bearish messages on StockTwits is extremely valuable as it allows the creation of a specific labeled vocabulary out of labeled messages. We are not aware of any other social media platform in finance offering this functionality. Second, text data from StockTwits is more reliable and less noisy than from general purpose platforms, such as Twitter, because messages focus on finance and economics matters only. Micro-bloggers have incentives to post valuable information in order to maintain or increase mentions and retweets, and thus have a greater share of voice in the forum (\citet{sprenger2014tweets}). On the other hand, StockTwits messages might be biased and subject to malicious users that try to manipulate the market. However, market manipulations likely happen only rarely as the SEC closely monitors potential influencers to prevent any market abuse. Third, extracting data from StockTwits is easy because of its API. StockTwits' API is designed to query the database to download messages via JSON requests. We provide a short tutorial in Appendix~\ref{app_tut}.

@@

Every StockTwits message includes eight features: (1) the reference ticker(s), (2) a timestamp, (3) a unique message identifier, (4) the body of the message, (5) the sentiment label (bearish, bullish, or none) entered by the user, (6) a unique identifier of the user who posted the message, (7) the number of messages published by the user who posted the message, and (8) the number of followers of the user who posted the message. Our sentiment analysis builds on the first five features. The last three provide additional information on the network structure, which we briefly discuss in Appendix \ref{app_userstats}.

@@

The left plot of Figure \ref{fig:data-stocktwits2} shows the top 30 most discussed tickers on StockTwits. SPY, a large ETF that tracks the S\&P 500 stock market index, is the most discussed ticker, followed by Apple and other big tickers. The messages about the 15 (30) biggest tickers represent 20\% (25\%) of the total number of messages, which indicates that users talk about a wide panel of tickers and not only big firms. The right graph shows a histogram of the number of messages per ticker. The x-axis is log-scaled because due to extreme values the distribution is highly skewed.

@@

The left plot of Figure \ref{fig:classif_time} shows the proportions of user sentiment-labeled messages across time. In the early years of the platform, most messages were unlabeled, presumably because users were not familiar with the sentiment label yet. Albeit the proportion of unlabeled messages monotonically declines over the years, almost 60\% of the more recent messages are still unlabeled. Overall, around 30 million messages are user-labeled and 60 million messages are unlabeled. We conjecture that by far not all unlabeled messages reflect market neutral opinions. Indeed, the right plot of Figure \ref{fig:classif_time} reveals that a substantial part of user-unlabeled messages is machine learning classified as bullish or bearish. Hence these user-unlabeled messages contain indeed market relevant information, which we are able to capture by our algorithm.

@@

    \mycaption{Message classification}{Left plot shows the proportions of user-labeled messages: bullish (green), bearish (red), and unlabeled (gray) across time. The right plot shows the proportions of machine learning classified messages: bullish (light green predicted, green user-labeled), bearish (light red predicted, red user-labeled), and neutral (gray) across time. Proportions are aggregated monthly.}
    \label{fig:classif_time}

@@

Among the user-labeled messages we find five times more bullish than bearish ones. This ratio indicates that investors are on average optimistic about the market, which is consistent with findings in the literature, e.g., \citet{renault2017intraday}. Such an imbalance is a well-known issue in machine learning classification as it creates a bias towards the over-represented class (see \citet{chawla2004special}). There are various ways to tackle class imbalance. An all-purpose standard approach in machine learning is to over-sample the minority class, which consists of randomly re-sampling from the minority class and thus artificially re-balance the class sizes in the data. We use a different approach, which is tailored for our setup. As not every message carries a substantial stock market relevant sentiment, we deviate from the bullish--bearish dichotomy. Instead, we create an auxiliary neutral sentiment class to account for messages that do not take a clear stand. See Appendix \ref{app_exampleclass} for examples of such neutral messages.

@@

The right plot of Figure \ref{fig:classif_time} shows the proportions of our machine learning classifications across time. Percentages of bearish (user-labeled and classified as bearish) and bullish (user-labeled and classified as bearish) messages are stable over time, suggesting that our classification method is robust. Even if most messages were not user-labeled in the early years of the platform, as seen in the left plot of Figure \ref{fig:classif_time}, we are now able to classify the sentiment of all messages in the sample, including a neutral class. Consistent with the over-representation of bullish messages observed in the user-labeled messages, there are substantially more messages classified as bullish than bearish. Examples of classified messages are given in Appendix \ref{app_exampleclass}.

@@

We next aggregate message sentiments on a daily ticker-level and across the market. Thereto, we denote $C_{i,t,j}=1,0,-1$ for bullish, neutral, bearish, respectively, the sentiment of the $j$th message about ticker $i$ on day $t$.\footnote{We follow the close-to-close convention. First, we remove all non-business days from the sample, whereby messages posted on non-business days count towards the next business day. ``Day $t$'' then stands for the time interval from 4pm on the previous business day $t-1$ to 4pm on business day $t$. This convention is consistent with the stock return data, which are close-to-close, and thus avoids any look-ahead bias of our sentiment polarity.} We follow \citet{ranco2015effects} and define the sentiment polarity of ticker $i$ on day $t$ as
\begin{equation}
    P_{i,t} = \dfrac{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} - \textbf{1}_{C_{i,t,j} = -1})}{\sum_{j=1}^{V_{i,t}} (\textbf{1}_{C_{i,t,j} = 1} + \textbf{1}_{C_{i,t,j} = -1})},
\end{equation}
where $V_{i,t}$ denotes the number of messages about ticker $i$ on day $t$.\footnote{If $V_{i,t}=0$ then we set $P_{i,t} =0$.}

@@

For the following time series analysis and event study we restrict our sample. There are two reasons for doing so. First, we keep computational cost at a reasonable level. Second, and more importantly, the time series of ticker-individual polarities exhibit spikes and are too noisy if the daily message volumes $V_{i,t}$ are too small. We therefore exclude from now on tickers whose median of daily message volume is less than 50. Our reduced sample contains 19 tickers, covering a range of security types, sectors, and market capitalization. We refer to Appendix \ref{app_coverage} for more details on the coverage, including summary statistics and the list of tickers covered.

@@

Event studies constitute a statistical method widely used in financial econometrics (see, e.g., \citet{mackinlay1997event}). In general, they are used to measure the effect of events on the market value of stocks. Well-known applications of event studies include the testing of various forms of the efficient market hypothesis (EMH) (see \citet{fama1969} and \citet{fama1991efficient}).

@@

Next we define the type of the event as either bullish, neutral or bearish. We use the abnormal polarity $AP_{i,t}$ of the event date to assess how on average investors perceive the event. Figure \ref{fig:abnpol_all} shows the distribution of abnormal polarities on event dates. 
\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{abnpol_distr_all.JPG}
    \mycaption{Empirical distribution of abnormal polarities on event dates}{The red dashed lines show the one-third and two-third percentiles.}
    \label{fig:abnpol_all}
\end{figure}
We chose to use the one-third (-0.03) and two-third percentile (0.07) of the distribution of abnormal polarities as thresholds for the type of the event. We define the type of the event for ticker $i$ at $t$ as
\begin{equation}
Type_{i,t}= \begin{cases}
      Bullish  & \text{if}\ \ AP_{i,t}> 0.07, \\
      Neutral  & \text{if}\ \ AP_{i,t} \in [-0.03, 0.07], \\
      Bearish  & \text{if}\ \ AP_{i,t} < -0.03. \\
\end{cases}    
\end{equation}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.75\textwidth]{events_across_time.JPG}
    \mycaption{Number of events of each type across time}{The green area shows bullish events, the gray area shows neutral events and the red area shows bearish events. Numbers are aggregated monthly.}
    \label{fig:events}
\end{figure}

As an illustration, Figure \ref{fig:activityaapl} shows for Apple the time series of message volume and the corresponding events. Between January 2011 and March 2020, our algorithm identified 73 events for Apple. 
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.75\textwidth]{activity_aapl.JPG}
    \mycaption{Daily message volume for Apple}{Events are days with an unusual high number of messages. The green upper-triangles show bullish events, gray circles are neutral events and red down-triangles represent bearish events.}
    \label{fig:activityaapl}
\end{figure}
What are these events? Remarkably, we capture a variety of corporate events and disclosures. Earning announcements constitute about half of the events. Other events include Apple Keynotes (presentations that Apple gives to the press, often presenting new products), or CEO letters addressed to investors. Table \ref{tab:event_aapl} lists a few examples.

@@

Left plot of Figure \ref{fig:caarcaap} shows the CAAR around the events. This plot is consistent with \citet{mackinlay1997event}. It shows that CAAR related to bearish (bullish) events exhibits a significant downward (upward) jump at the event date, respectively. These jumps are followed by a flat CAAR during the 20 days after the event. Interestingly, there is a systematic shift in the CAAR already 1 day before the event. However, this shift is relatively small compared to the jump on the event day: one day before the event, the bullish (bearish) CAAR equals 0.019 (-0.023). The CAAR related to the neutral events exhibits a slight upward shift around the event date but it fades away after a few days. The CAAR related to bearish events shifts already a few days before the event but this shift is not statistically significant. This is in line with Figure \ref{fig:boxplotcaar}, which shows that the CAR distributions prior to the events are not significantly different from zero. This is confirmed by the Mann-Whitney U-tests shown in Table \ref{utest_res}. CAR has no predictive power on the type of the event: five days before an event, the median of the CAR distribution of the bullish events is not statistically different from the median of the neutral events. The same holds for the bearish events.

@@

%\clearpage
\subsection{Abnormal Polarity}
How does sentiment polarity behave around events? Similar to the above, we use as benchmark model a one-year rolling window regression of the daily polarity of ticker $i$ on the daily market polarity defined in \eqref{marketpoleq}, 
\begin{equation}
    P_{i,t}= \alpha_i^P  + \beta_i^P \cdot P_t^M  + \epsilon_{i,t},
\end{equation}
which gives the one-year rolling estimates $\hat{\alpha}_i^P$ and $\hat{\beta}_i^P$. This implies the abnormal polarity 
\begin{equation}
      AP_{i,t} = P_{i,t} - \left(\hat{\alpha_i}^P + \hat{\beta_i}^P \cdot P_t^M\right).
\end{equation}
We define the cumulative abnormal polarity (CAP) around a ticker $i$ event $\tau$ as 
\begin{equation}\label{eq_CAPtaut}
    CAP_i(\tau,t) = \sum_{s=- 20}^{t} AP_{i,\tau + s},
\end{equation}     
and the cumulative average abnormal polarities (CAAP) across all $N=1131$ events as 
\begin{equation}
       CAAP(t) = \dfrac{1}{N} \sum_{j=1}^N CAP_{i_j}(\tau_j,t).
\end{equation}

The right plot of Figure \ref{fig:caarcaap} shows the CAAP around the events. There are two main findings. First, in contrast to CAAR, the CAAP for bullish and bearish events is not constant after the event date, suggesting that users' sentiments about stocks tend to be biased towards recent past events. A possible explanation is that users might still post bullish (bearish) messages about a bullish (bearish) event during several days after the event. This is in contrast to the returns that immediately normalize after the event. Second, and more interestingly, the CAAP for bullish and bearish events shifts several days earlier than the CAAR. This indicates that investors are on average able to anticipate the type of an event in the near future. However, this sentiment only manifests through the social media, but not through abnormal returns.

@@

Based on these signals, we now construct reset-CAP-sorted portfolios. Formally, we define the ticker sets $I^{bull}_t=\{ i\mid CAP_{i,t}^{(R)} > U_t(x)\}$ and $I^{bear}_t=\{ i\mid CAP_{i,t}^{(R)} < L_t(x)\}$. At the close of any day $t$, we form the equally weighted bullish (bearish) portfolio consisting of tickers in $I^{bull}_t$ (in $I^{bear}_t$), and realize the 1-day returns. If any of the index sets is empty, we set the corresponding return to zero. The top-left plot of Figure \ref{fig:pf} shows the cumulative log returns of bullish and bearish portfolios as well as the S\&P500. Overall, the portfolio performance is consistent with our approach: the bullish (bearish) portfolio outperforms (under-performs) the market. Remarkably, the upward (downward) steps suggest that our portfolio strategy succeeds to take the right positions just before an event. The remaining plots of Figure \ref{fig:pf} show the number of positions across time of our portfolios. Most of the returns are earned with portfolios consisting of very few tickers. This is a result of our market timing and stock picking strategy: we only invest in the top/bottom percentiles of CAP, whenever our signal is strong enough.
